---
title: "P02- SNA_FinalProject - 2022"
output:
  html_document: default
  pdf_document: default
Author: Elham Babaei
---

```{r setup}
library(statnet)

#library("network") 
#library("sna")     # sns allows to do descriptive analysis in network
library("igraph")       # use igraph for visualization 

install.packages("CliquePercolation")
library(CliquePercolation)

#if(!require(weatherData)) install.packages("weatherData",repos = "http://cran.us.r-project.org")
#install.packages("weatherData",repos = "http://cran.us.r-project.org")
```

In P02 file, we have 301 authors and 4766 events. The incidence matrix shows a two mode network, author-paper relationship.

There are some repeated rows in the main dataset. We have to make sure all the values in incidence matrix are 0 and 1. If an author has participated in a paper the value is 1 otherwise is 0.
```{r}
# Incidence matrix - two-mode
incmat <- read.csv("P02_IncMat.csv",header = T, row.names = 1)
incmat <- as.matrix(incmat)
#incmat

class(incmat)
typeof(incmat)
dim(incmat)    # authors - papers
incmat[incmat>1] <- 1          
min(incmat)
max(incmat)

#gplot(IncMat, gmode="twomode",displaylabels = F)
```

We change the diagonal of adjacency matrix to values zero to avoid circle edges to each node itself because in this case We cannot obtain nice visualizations. 

```{r}
#Adjacency matrix - one-mode
adjmat <- incmat %*% t(incmat)    # nodes ( author - author)
diag(adjmat) <- 0
class(adjmat)
typeof(adjmat)
dim(adjmat)
min(adjmat)
median(adjmat)
max(adjmat)    # maximum weight between pair of authors ( some authors have participated in 172 common events)

#adjmat
```

The attributes extracted from the dataset are the following:

Role, University, Localization, Faculty, Gender, Academic_Age, Age   
```{r}
# Import vertex attributes (this can be left as dataframe)
nodeInfo <- read.csv("P02_NodeAttr.csv",header=TRUE,stringsAsFactors=FALSE, sep=",")
nodeInfo
```

```{r}
summary(nodeInfo)
```

We create the network from the adjacency matrix.

```{r}
# network
net <- network(adjmat,directed=FALSE) 

# Add nodal attributes to a network object
net %v% "Role" <- nodeInfo$Role
net %v% "University" <- nodeInfo$University
net %v% "Localization"  <- nodeInfo$Localization
net %v% "Faculty"  <- nodeInfo$Faculty
net %v% "Gender" <- nodeInfo$Gender
net %v% "Academic_Age" <- nodeInfo$Academic_Age
net %v% "Age"  <- nodeInfo$Age

net         # Get a quick description of the data
```

```{r}
network::list.vertex.attributes(net) 
net %v% "vertex.names"          
```


# Descriptive Analysis and Visualization 

We use either `network` and `igraph` package for this purpose.

```{r}
is.network(net)
class(net)
network.size(net)
sna::is.connected(net)   #False
isolates(net)            #shows the position of isolated nodes
length(isolates(net))    # shows the number of isolated nodes
```

The density of the network is very small
```{r}
gden(net, mode="graph")    #density - sna
```

```{r}
# Dyadic reciprocity   this graph is not directed

#grecip(net, measure="dyadic.nonnull") 
#grecip(net, measure="dyadic")
#grecip(net, measure="edgewise")
```


```{r}
# Transitivity
trans <- gtrans(net, mode="graph") 
trans
```

```{r}
# Degree
deg <- sna::degree(net, gmode = "graph")    #shows the degree of all nodes
deg
max(deg)

```

```{r}
hist(deg, xlab="degree", main="P02- Degree distribution", prob=F, col = "light blue")
```


```{r}
#Betweenness
bet <- sna::betweenness(net, gmode="graph") 
bet       #not normalized

betdf <- data.frame(net %v% "vertex.name", bet) 
betdf[order(betdf$bet, decreasing = TRUE), c(1,2)]

gplot(net, vertex.cex=sqrt(bet)/25, gmode="graph", displaylabels = F, 
      label.cex = 0.8, main="P02- Network \n Node size: sqrt(betweennes)/25") 
```

Since the graph is not connected, closeness centrality cannot be defined.
```{r}
# Closeness
clo <- sna::closeness(net) 
#clo
#hist(clo)

```



```{r}
# Eigenvector centrality
ev <- sna::evcent(net) 
ev
hist(ev, main = "P02- Eigenvector Centrality", col="light green")

```

```{r}
#Subnetwork censuses
dc <- sna::dyad.census(net)
dc
tc <- sna::triad.census(net, mode = 'graph')
tc
barplot(tc, names.arg=colnames(tc), las=2, col=7, main = "P02- Triadic census")
#tc[-1]  # all the values except the first one
#barplot(tc[-1]/sum(sna::triad.census(net)[-1]), names.arg=colnames(tc)[-1], las=2)
```
``` {r}
# Kpath and kcycle
kp <- kpath.census(net, mode="graph", maxlen=5, tabulate.by.vertex=T) # Count paths of length <=5
#kp

kc <- kcycle.census(net, maxlen = 5, mode = "graph", tabulate.by.vertex = T) # Count cycles of length <=5
#kc
```

Since the graph is not directed, strong and weak connected components are the same.

There are 58 connected components.
``` {r}
#Components 
comp <- sna::components(net) # the number of components
comp

cl <- component.largest(net)  # true if the vertex belongs to largest component, false otherwise

gplot(net[cl,cl], boxed.lab=FALSE, label.cex=0.5,label.col=4, displaylabels = F, gmode = "graph",
      label=network.vertex.names(net)[cl], main="P02- The largest connected component")

``` 

```{r}
#Cliques census
cc<-clique.census(net,clique.comembership="bysize")
#cc
```

# Descriptive Analysis and Visualization using igraph package

The values in the adjecency matrix are considered as edge weight 
```{r}
# create a graph (igraph object)
options(warn = -1)
mygraph <- graph_from_adjacency_matrix(adjmatrix = adjmat, mode="undirected", weighted = TRUE)

mygraph <- set_vertex_attr(mygraph,name = "Role", value = as.character(nodeInfo$Role))
mygraph <- set_vertex_attr(mygraph,"University", value = as.character(nodeInfo$University))
mygraph <- set_vertex_attr(mygraph,"Localization", value = as.character(nodeInfo$Localization))
mygraph <- set_vertex_attr(mygraph,"Faculty", value = as.character(nodeInfo$Faculty))
mygraph <- set_vertex_attr(mygraph,"Gender", value = as.character(nodeInfo$Gender))
mygraph <- set_vertex_attr(mygraph,"Academic_Age", value = as.integer(nodeInfo$Academic_Age))
mygraph <- set_vertex_attr(mygraph,"Age", value = as.integer(nodeInfo$Age))

mygraph
```

The distribution of edge weights is very skewed and shows that the pair of authors usually contribute in less than 5 papers together. The maximum weight is 49.
```{r}
max(E(mygraph)$weight)    # is the same max value in adj matrix . correct
min(E(mygraph)$weight)
#V(mygraph)$name

class(E(mygraph)$weight)
#E(mygraph)$weight

hist(E(mygraph)$weight, col="lightblue", xlab="Edge weight", ylab="Frequency", 
     main="P02- Edge weight Distribution")

```

```{r}
class(mygraph)
vcount(mygraph)
ecount(mygraph)
edge_density(mygraph)
is.connected(mygraph)

plot(mygraph,vertex.size=4 ,vertex.label=NA, vertex.label.cex=0.2, edge.arrow.size=0.4,
     main="P02- Graph")

```

```{r}
igraph::dyad.census(mygraph)

igraph::triad.census(mygraph)  #Undirected graph
reciprocity(mygraph, mode="ratio") 

```

Here we look at the neighbors of each node having the distance 1 or 2 from that node.
We also show the graph for the node `Adelchi Azzalini`.
```{r}
# neighborhood
k.nbhds <- graph.neighborhood(mygraph, order=1) 
sapply(k.nbhds, vcount)
k.1 <- k.nbhds[[1]]   # the subgraph created by neighbors of node number 1 with distance equal to 1
edge_density(k.1)


k.nbhds2 <- graph.neighborhood(mygraph, order=2) 
sapply(k.nbhds2, vcount)
k.1.2 <- k.nbhds2[[1]]   # the subgraph created by neighbors of node number 1 with distance equal to 2
edge_density(k.1.2)

edge_density(mygraph)

plot(k.1, vertex.label.cex=0.6, vertex.cex=0.5 , 
     main=" P02- Neighborhood of order 1 for 'Adelchi Azzalini' ")
plot(k.1.2, vertex.label.cex=0.5, vertex.cex=0.1, 
     main=" P02- Neighborhood of order 2 for 'Adelchi Azzalini' ")

```

```{r}
# degree dist in igraph
d.mygraph <- degree(mygraph)
#d.mygraph
dd.mygraph <- degree.distribution(mygraph)
#dd.mygraph

hist(degree(mygraph), col="lightblue", xlim=c(0,12), xlab="Vertex Degree", ylab="Frequency", 
     main="P02- Degree Distribution") 

```
Vertex strength for each vertex is the sum of edge weights of the adjacent edges for that vertex.
```{r}
#Vertex strength
hist(graph.strength(mygraph), col="pink", xlab="Vertex Strength", ylab="Frequency",
     main="P02- Vertex strength distribution") 
```
Edge betweennes is defined by the number of geodesic distances (shortest paths) going through that edge.

The result shows that the statistician `Francesco Bartolucci` plays the most important role in terms of edge betweeness in the graph. This shows that he has many connection with other authors. However, it this was more obvious for P01 dataset
```{r}
# Edge betweennes 
eb <- edge.betweenness(mygraph)
#eb
E(mygraph)[order(eb, decreasing=T)[1:5]] # the first five edges with highest edge betweeness

```

There are 1535 cliques of sizes from 1 to maximum 9.
```{r}
#cliques
cliques.mygraph <- cliques(mygraph)

length((cliques.mygraph))
#shows the number of cliques of different sizes 
table(sapply(cliques(mygraph), length)) 


```

We look at clique of size 9.

```{r}
which(sapply(cliques(mygraph), length)==9)  # which cliques of order 9
cliques.mygraph[1286]
```

```{r}
# the induced subgraph created by the nodes of clique of order 9
plot(induced.subgraph(mygraph, cliques.mygraph[[1286]]), vertex.cex=2, vertex.label.cex=0.8,
     main="P02- The largest clique of size 9")  

```

Cliques of larger sizes necessarily include cliques of smaller sizes.
For instance cliques of order 4 are included in cliques of order 5. so they are not counted in maximal cliques
```{r}
# the number of maximal cliques of different sizes 
print("Maximal cliques of different sizes: ")
table(sapply(maximal.cliques(mygraph), length)) 
```

We look at the induced subgraph created by neighbors of distance 1 for the node professor `Nicola Torelli` and the density of the resulted subgraph which is higher than the density of the whole graph.
```{r}
#Ego-density (or local density)
ego.instr <- induced.subgraph(mygraph, neighborhood(mygraph, order= 1, nodes= 230)[[1]]) 
#neighborhood of node number 230 (or we can use the label of node, node="Nicola TORELLI") 
#order is the same geodesic distance                                                                  
edge_density(mygraph)
edge_density(ego.instr)

plot(ego.instr, vertex.cex=0.3, vertex.label.cex=0.8, main="P02- Induced subgraph for 'Nicola Torelli'")   
```
The authors `Nicola TORELLI`, `Matilde TREVISANI`,and `Francesco PAULI`, are from the University of Trieste and `Fortunato Pesarin` is from the University of Padova.

Then we look at the local transitivity for these authors.
```{r}
#transitivity
transitivity(mygraph)   # the ratio of the count of triangles and connected triples in the graph.

#The local transitivity of a vertex is the ratio of the count of triangles connected to the vertex 
#and the triples centered on the vertex
transitivity(mygraph, "local", vids=c(230,210, 114, 97))
transitivity(mygraph, "local", vids=c("Nicola TORELLI", "Matilde TREVISANI",
                                      "Francesco PAULI", "Fortunato PESARIN"))

```

```{r}
# path length
print("Average path length:")
average.path.length(mygraph)
print("Diameter:")
diameter(mygraph)            # the length of the longest geodesic distance.
```

The number of connected components is the same 58 that we obtained using the Network package.

```{r}
#components

table(sapply(decompose.graph(mygraph), vcount))  

comp <- igraph::components(mygraph)

#groups(comp)

#component_distribution(mygraph, cumulative = FALSE, mul.size = FALSE)

allcomps <- decompose(mygraph)

gc <- allcomps[[1]] #giant component
print(" Giant component:")
gc
#vcount(gc)
#ecount(gc)

plot(gc,vertex.label="",  vertex.size= 3, main="P01- The giant component") 
plot(allcomps[[4]])


```


```{r}
#cut vertices in the giant component
gc.cut.vertices <- articulation.points(gc)

print("The number and proportion of cut vertices in giant component:")
length(gc.cut.vertices)
length(gc.cut.vertices)/vcount(gc) # the proportion of cut vertices 
                                                # if this value is high, the network is vulnerable

```

```{r}
#cut vertices in the graph
mygraph.cut.vertices <- articulation.points(mygraph)

print("The number and proportion of cut vertices in the whole graph:")
length(mygraph.cut.vertices)
length(mygraph.cut.vertices)/vcount(mygraph) # the proportion of cut vertices 
                                                # if this value is high, the network is vulnerable
```


#Visualization


```{r}
# looking at some useful vertex attributes for visualization
table(V(mygraph)$Gender)
table(V(mygraph)$Localization)
table(V(mygraph)$Role)

table(V(mygraph)$Faculty)
table(V(mygraph)$Academic_Age)
table(V(mygraph)$Age)
table(V(mygraph)$University)
```

```{r}
gplot(incmat, gmode="twomode",displaylabels = F, label.cex = 0.7)

gplot(adjmat, gmode="graph",displaylabels = F, label.cex = 0.7, edge.lwd=adjmat, main="P02-Network") 
## the lines are proportional to edg.lwd= adjmat as weights
```
First of all we visualize the network using different layouts and the strength of the nodes as the node size.
Then we pick the best visualization and we try other measures for node size on it.

We also consider orange color for female and light blue color for male authors. 
The edges between two female authors are pink, between two male authors are blue and between a male and a female author dark gray.

Also, the width of edges is specified by their weight divided by 10.

```{r}
set.seed(42)

E(mygraph)$width <- E(mygraph)$weight/10

#V(mygraph)[V(mygraph)$Role=="PA"]$color <-1
#V(mygraph)[V(mygraph)$Role=="PO"]$color <-2
#V(mygraph)[V(mygraph)$Role=="RU"]$color <-3

#V(mygraph)[V(mygraph)$Gender=="F"]$shape <- "circle"
V(mygraph)[V(mygraph)$Gender=="F"]$color <- "orange"
#V(mygraph)[V(mygraph)$Gender=="M"]$shape <- "square"
V(mygraph)[V(mygraph)$Gender=="M"]$color <- "lightblue"

V(mygraph)$size <- sqrt(graph.strength(mygraph))   # strength of the node as the nodes size 
#V(mygraph)$size2 <- V(mygraph)$size * .5

female <- V(mygraph)[Gender=="F"]
male  <- V(mygraph)[Gender=="M"]
E(mygraph)[ female %--% male ]$color <- "black"          
E(mygraph)[ female %--% female ]$color <- "red"    
E(mygraph)[ male %--% male ]$color <- "blue" 

#igraph.options(vertex.size=6, vertex.label=NA, vertex.label.dist=1, vertex.label.cex=0.2, edge.color="ligh tblue", edge.arrow.size=0.4)

```

 
```{r}
# random layout
plot(mygraph, layout=layout_randomly, vertex.label=NA)
title(" P02- Random layout\n Node size = node strength")
legend('bottomleft', fill=c("orange","lightblue", "red", "blue", "black"),
       legend=c("Femal (F)","Male (M)", "F-F co-outhorship",
                "M-M co-outhorship",
                "F-M co-outhorship"),cex=0.75)
```

```{r}
edge_weight <- E(mygraph)$weight
print("P02-The first five pair of authors with highets co-authorship")
E(mygraph)[order(edge_weight, decreasing=T)[1:5]]

```

```{r}
# nicely layout
plot(mygraph, layout=layout_nicely, vertex.label=NA)
title(" P02- Nicely layout\n Node size = node strength")
legend('bottomleft', fill=c("orange","lightblue", "red", "blue", "black"),
       legend=c("Femal (F)","Male (M)", "F-F co-outhorship",
                "M-M co-outhorship",
                "F-M co-outhorship"),cex=0.75)
```


```{r}
# circular layout
plot(mygraph, layout=layout.circle, vertex.label=NA)
title(" P02- Circular layout\n Node size = node strength", cex.main=0.8)
legend('bottomleft', fill=c("orange","lightblue", "red", "blue", "black"),
       legend=c("Femal (F)","Male (M)", "F-F co-outhorship",
                "M-M co-outhorship",
                "F-M co-outhorship"),cex=0.75)
```

```{r}
# Fruchterman & Reingold layout
plot(mygraph, layout=layout_with_fr, vertex.label=NA)
title(" P02- Fruchterman & Reingold layout\n Node size = node strength")
legend('bottomleft', fill=c("orange","lightblue", "red", "blue", "black"),
       legend=c("Femal (F)","Male (M)", "F-F co-outhorship",
                "M-M co-outhorship",
                "F-M co-outhorship"),cex=0.75)
```

```{r}
# MDS layout
plot(mygraph, layout=layout_with_mds(mygraph), vertex.label=NA)
title(" P02- MDS layout \n Node size = node strength")
legend('bottomleft', fill=c("orange","lightblue", "red", "blue", "black"),
       legend=c("Femal (F)","Male (M)", "F-F co-outhorship",
                "M-M co-outhorship",
                "F-M co-outhorship"),cex=0.75)
```

```{r}
# Kamada-Kawai layout
plot(mygraph, layout=layout_with_kk, vertex.label=NA)
title(" P02- Kamada-Kawai layout \n Node size = node strength")
legend('bottomleft', fill=c("orange","lightblue", "red", "blue", "black"),
       legend=c("Femal (F)","Male (M)", "F-F co-outhorship",
                "M-M co-outhorship",
                "F-M co-outhorship"),cex=0.75)
```
It seems the best visualization is the one using the layout `Fruchterman & Reingold`.

We try it with other measures of node size.
```{r}
#Compute node degree and use it to set node size in the graph
deg_g <- igraph::degree(mygraph, mode = "all", normalized = T)
V(mygraph)$size <- deg_g*100     # because these values are small
#V(mygraph)$name <- 1:vcount(mygraph)
V(mygraph)$label.color <- "blue"
plot(mygraph, layout=layout_with_fr, vertex.label = " ", vertex.label.dist=0, vertex.label.cex=0.6) 
title(main="P02- Fruchterman & Reingold layout layout.\n Node size = node degree centrality", cex.main=1)
legend('bottomleft', fill=c("orange","lightblue", "red", "blue", "black"),
       legend=c("Femal (F)","Male (M)", "F-F co-outhorship",
                "M-M co-outhorship",
                "F-M co-outhorship"),cex=0.75)
```

Closeness cannot be used since the graph in not connected.
```{r}
# Compute node betweenness and use it to set node size

betw_g <- igraph::betweenness(mygraph, directed = F, normalized = T)
#betw_g
V(mygraph)$size <- betw_g*100
V(mygraph)$label.color <- "blue"
plot(mygraph, layout=layout_with_fr, vertex.label = " ",  vertex.label.dist=0, vertex.label.cex=0.6) 
title(main="P02- Fruchterman & Reingold layout\n Node size = betweenness centrality", cex.main=1)
legend('bottomleft', fill=c("orange","lightblue", "red", "blue", "black"),
       legend=c("Femal (F)","Male (M)", "F-F co-outhorship",
                "M-M co-outhorship",
                "F-M co-outhorship"),cex=0.75)
```
The above visualization shows the authors with highest betweenness are females.
```{r}
#Compute node eigenvector and use it to set node size

eigen_g <- eigen_centrality(mygraph)
V(mygraph)$size <- eigen_g$vector*10
V(mygraph)$label.color <- "blue"
plot(mygraph, layout=layout_with_fr, vertex.label = "",  vertex.label.dist=0, vertex.label.cex=0.6) 
title(main="P02- Fruchterman & Reingold layout layout.\n Node size = eigenvector centrality", cex.main=1)
legend('bottomleft', fill=c("orange","lightblue", "red", "blue", "black"),
       legend=c("Femal (F)","Male (M)", "F-F co-outhorship",
                "M-M co-outhorship",
                "F-M co-outhorship"),cex=0.75)
```

```{r}

mygraph.adiacency <- graph_from_adjacency_matrix(adjmat, mode="undirected")
mygraphm <- get.adjacency(mygraph.adiacency, sparse=F)     # is the same adj matrix

class(mygraph.adiacency) # is a graph object
edge_density(mygraph.adiacency)

# Degree distribution
par(mar=c(4,4,4,4)) ### set margins
dd <- degree.distribution(mygraph.adiacency, cumulative=T, mode="all")
plot(dd, pch=19, cex=0.8, col="orange", xlab="Degree", ylab="Cumulative", main = "P02- Cumulative degree distribution")
```

```{r}
# Heatmap of the network matrix - sociomatrix representation
#### reordering adiacency matrix according to clustering - geodesic distances 
geomygraph<-distances(mygraph.adiacency, v = V(mygraph.adiacency), to = V(mygraph.adiacency), mode ="all", weights = NULL, 
                  algorithm = "automatic")
geomygraph[geomygraph==Inf]<-15
geomygraph<-as.dist(geomygraph)
hclust_geomygraph<-hclust(geomygraph, method = "complete", members = NULL)  #heirarchical clustering
heatmap(mygraphm,Rowv=hclust_geomygraph$order,Colv="Rowv", scale="none")
title(main="P02- heatmap plot ", cex.main=0.8)
```

```{r}
###heatmap by gender
palf <- colorRampPalette(c("lightblue", "red")) 
heatmap(mygraphm[order(V(mygraph)$Gender), ], col = palf(10), labRow= rownames(mygraph), labCol= colnames(mygraph))
title(main="P02- heatmap plot by gender ", cex.main=0.8)
```


```{r}

### Circos graphs
library(circlize)

colnames(mygraphm) <- V(mygraph.adiacency)$name
rownames(mygraphm) <- V(mygraph.adiacency)$name

circos.clear()
par(cex=0.5)
circos.par( start.degree = 90, clock.wise = T)
chordDiagram(mygraphm, symmetric=T, annotationTrack = c("name","grid"))       
# the larger the slice in circle graph the larger the degree of the node
```


```{r}

### change the orientation of node label
### pre-allocating an empty track and customizing the labels in it later
chordDiagram(mygraphm, symmetric=T, annotationTrack = "grid", 
preAllocateTracks = list(track.height = max(strwidth(unlist(dimnames(mygraphm))))))

circos.track(track.index = 1, panel.fun = function(x, y) {
  circos.text(CELL_META$xcenter, CELL_META$ylim[1], CELL_META$sector.index, 
              facing = "clockwise", niceFacing = TRUE, adj = c(0, 0.5))
}, bg.border = NA) # here set bg.border to NA is important
```


```{r}
# now, the image with rotated labels
chordDiagram(mygraphm, annotationTrack = "grid", preAllocateTracks = 1)
circos.trackPlotRegion(track.index = 1, panel.fun = function(x, y) {
  xlim = get.cell.meta.data("xlim")
  ylim = get.cell.meta.data("ylim")
  sector.name = get.cell.meta.data("sector.index")
  circos.text(mean(xlim), ylim[1] + .1, sector.name, facing = "clockwise", niceFacing = TRUE, adj = c(0, 0.5))
  circos.axis(h = "top", labels.cex = 0.5, major.tick.percentage = 0.2, sector.index = sector.name, track.index = 2)
}, bg.border = NA)
```

# Community Detection

The following chunk shows the number of nodes in each of 58 connected components.

```{r}
# membership to components (we have 45 components)

mygraph.scc <- clusters(mygraph) # Extract the maximal (weakly or strongly) connected components of a graph

#table(mygraph.scc$csize)

#mygraph.scc$membership

table(mygraph.scc$membership)

scc <- induced.subgraph(mygraph, vids= mygraph.scc$membership==28)

plot(scc, vertex.size=2 )
```


```{r}
#K-core decomposition allows us to identify the core and the periphery of the network. A k-core is a maximal subnet of a network such that all nodes have at least degree K.

# compute coreness
kcores <- coreness(mygraph)
coreness <- kcores
str(coreness(mygraph))


V(mygraph)$core <- coreness(mygraph, "all") # assign as node attribute the k-core to which nodes belong
```


```{r}
# create layout

CorenessLayout <- function(g) {
  coreness <- graph.coreness(g);
  xy <- array(NA, dim=c(length(coreness), 2));
  
  shells <- sort(unique(coreness));
  for(shell in shells) {
    v <- 1 - ((shell-1) / max(shells));
    nodes_in_shell <- sum(coreness==shell);
    angles <- seq(0,360,(360/nodes_in_shell));
    angles <- angles[-length(angles)]; # remove last element
    xy[coreness==shell, 1] <- sin(angles) * v;
    xy[coreness==shell, 2] <- cos(angles) * v;
  }
  return(xy);
}



ll <- CorenessLayout(mygraph)

# plot
plot(mygraph, layout=ll, vertex.size=6, vertex.label=" ", vertex.label.cex=0.8, main='P02-Coreness')

```

```{r}
# if you just need to know the vertices and not to build the subgraph 

maxCoreness <- max(coreness)
maxCoreness     # so k-core is actually 8-core
verticesHavingMaxCoreness <- which(coreness == 8) 
kcore <- induced.subgraph(graph=mygraph,vids=verticesHavingMaxCoreness)

plot(kcore,  vertex.label=igraph::get.vertex.attribute(kcore,name='vert.names',index=V(kcore)), 
     vertex.size=5,vertex.cex=2, vertex.label.cex=0.8, main="P02- Vertices having coreness 8")

```
With Girvan-newman algorithm we obtain 73 clusters.

Cluster number 3 has the maximum number of members which is 42
```{r}
# Girvan-newman algorithm
options(warn = -1)

gn.comm <- cluster_edge_betweenness(mygraph) 

table(gn.comm$membership)
max(table(gn.comm$membership))
```

```{r}
is_hierarchical(gn.comm)
plot_dendrogram(gn.comm)
```

```{r}
# visualizing the result of dividing the network into communities
colors <- rainbow(max(membership(gn.comm)))
plot(gn.comm, mygraph, vertex.size = 6, vertex.color=colors[membership(gn.comm)], 
     vertex.label = NA, edge.width = 1, main= "P02- Clusters from Girvan-Newman algorithm")

```
Modularity using Girvan-Newman algorithm is high and positive. so it is good.
```{r}
# computing modularity
length(unique(membership(gn.comm))) # the number of clusters
mod.gn <- modularity(mygraph,membership(gn.comm)) 
mod.gn
```

```{r}
# compute the intra-cluster density
print("Intra cluster density for Girvan-Newman algorithm algorithm:")
sapply(unique(membership(gn.comm)), function(gg) {
  subg1<-induced.subgraph(mygraph, which(membership(gn.comm)==gg)) #membership id differs for each cluster
  edge_density(subg1)
})

edge_density(mygraph) #overall density is not high
```

```{r}
#get all combinations of communities and get inter-cluster edges

cs <- data.frame(combn(unique(membership(gn.comm)),2))
cx <- sapply(cs, function(x) {
  es<-E(mygraph)[V(mygraph)[membership(gn.comm)==x[1]] %--% 
                  V(mygraph)[membership(gn.comm)==x[2]]]    
  length(es)
})
cbind(t(cs),inter.edges=cx)  # shows the number of edges between pair of communities. 
```

```{r}
#evaluating cluster homogeneity

#Gini index (complementary). The larger its value the larger cluster homogeneity according to a given actor attribute
gini <- function(x){
  f <- table(x)/length(x)
  sum(f^2)
}
```


```{r}
#attach community labels as vertex attribute
V(mygraph)$GN.cluster <- membership(gn.comm)

gini(V(mygraph)$Role[V(mygraph)$GN.cluster==3])   # 3 is the largest cluster. output 1 means everyone in cluster number 24 belongs to the same community
```

gini index for cluster number 3 is not that high.

```{r}
# Louvain algorithm (modularity optimization)

louv.clu <- cluster_louvain(mygraph)

table(louv.clu$membership)
max(table(louv.clu$membership))

```
```{r}
# visualizing the result of dividing the network into communities
colors <- rainbow(max(membership(louv.clu)))
plot(louv.clu, mygraph, vertex.size = 6, vertex.color=colors[membership(louv.clu)], vertex.label = NA, edge.width = 1, main= "P02- Clusters from Louvain algorithm")
```

```{r}
length(unique(membership(louv.clu)))     
modularity(mygraph,membership(louv.clu)) 
```

```{r}

#attach community labels as vertex attribute
V(mygraph)$LOUV.cluster <- membership(louv.clu)

gini(V(mygraph)$Role[V(mygraph)$LOUV.cluster==1])   # 1 is the largest cluster. output 1 means everyone in cluster number 24 belongs to the same community

```

```{r}
# compute the intra-cluster density
print("Intra cluster density for Louvian algorithm:")
sapply(unique(membership(louv.clu)), function(gg) {
  subg1<-induced.subgraph(mygraph, which(membership(louv.clu)==gg)) #membership id differs for each cluster
  edge_density(subg1)
})

edge_density(mygraph) #overall density is not high
```



Girvan Newman algorithm gives 73 clusters with modularity 0.71, while Louvain algorithm provides the same number of clusters but with modularity 0.69.

So Girvan Newman algorithm is better in this case because of higher modularity. 

```{r}
# Clique percolation algorithm

# weighted density
sum(E(mygraph)$weight)/(vcount(mygraph)*(vcount(mygraph)-1)/2)

# unweighted density
length(E(mygraph)$weight)/(vcount(mygraph)*(vcount(mygraph)-1)/2)

```

Considering k=7 for Clique Percolation algorithm, there is one node shared between two cliques.
```{r}
options(warn = -1)

cp.comm_w <-  cpAlgorithm(get.adjacency(mygraph, attr = "weight", sparse = F), k=7, method = "weighted", I=0.5) 
#I: intensity threshold parameter
#k is started from 3. 2 is not aallowed

#cp.comm_w$list.of.communities.labels
cp.comm_w$list.of.communities.numbers

#cp.comm_w$shared.nodes.labels
cp.comm_w$shared.nodes.numbers
```


# ERGM models


```{r}
# choosing a subgraph for ergm and dychotomising the edge weights
ergm_graph = mygraph

E(ergm_graph)$weight[E(ergm_graph)$weight < 10]=0
E(ergm_graph)$weight[E(ergm_graph)$weight >0]=1
#E(ergm_graph)$weight

ergm_graph$vertexIDs <- c(1:vcount(ergm_graph))

ergm_induced_graph = induced.subgraph(ergm_graph, 
                                      vids = ergm_graph$vertexIDs[c(250:300)] )

#ergm_induced_graph = subgraph.edges(ergm_graph, E(ergm_graph)[which(E(ergm_graph)$weight>0)])

ergm_induced_graph
vcount(ergm_induced_graph)
ecount(ergm_induced_graph)
plot(ergm_induced_graph, vertex.label=NA, vertex.size=5, main="P02- Sub-graph for ERGM")
```
```{r}
# creating the network object for ergm

ergm_induced_graph.m <- get.adjacency(ergm_induced_graph, sparse=F)     # is the same adj matrix

ergm_induced_net <- as.network(ergm_induced_graph.m, directed=F)

# Add nodal attributes to a network object
ergm_induced_net %v% "Role" <- get.vertex.attribute(ergm_induced_graph)$Role
ergm_induced_net %v% "University" <- get.vertex.attribute(ergm_induced_graph)$University
ergm_induced_net %v% "Localization"  <- get.vertex.attribute(ergm_induced_graph)$Localization
ergm_induced_net %v% "Faculty"  <- get.vertex.attribute(ergm_induced_graph)$Faculty
ergm_induced_net %v% "Gender" <- get.vertex.attribute(ergm_induced_graph)$Gender
ergm_induced_net %v% "Academic_Age" <- get.vertex.attribute(ergm_induced_graph)$Academic_Age
ergm_induced_net %v% "Age"  <- get.vertex.attribute(ergm_induced_graph)$Age

ergm_induced_net
plot(ergm_induced_net, main="P02- Sub-network for ERGM")
```

The input of ergm model must be a network object. 

So here we use the network that we created from the data. But since it is too large for ergm models, we choose an induced subgraph of it which considers only the cental zone in Italy.

```{r}

#network::list.vertex.attributes(net) 
#net %v% "vertex.names"   


#inducednet <- get.inducedSubgraph(net,v= d)
#inducednet <- get.inducedSubgraph(net,v= (which(net%v%"Localization"=="North")))

#inducednet <- get.inducedSubgraph(net,v= (which(net%v%"Localization"=="North")))

#inducednet <- get.inducedSubgraph(net,v= (which(net%v%"Academic_Age" <10 & net%v%"Gender"=="M")))
#inducednet <- get.inducedSubgraph(net,v= (which(net%v%"Academic_Age" <=7 )))
#inducednet <- get.inducedSubgraph(net,v= (which(net%v%"University"=="Univ. TRIESTE" ))) 

#inducednet
#plot(inducednet)
```


```{r}
#Let's start with baseline model (ER model)
model.00 <- ergm(ergm_induced_net~edges) # fit model
summary(model.00) # look in more depth
#coef(model.01)

```

The probability of activating a tie randomly is equal to the density of the network.
```{r}
# the probability of activating a tie randomly
print("The probability of activating a tie randomly:")
plogis(coef(model.00)[1])
print("Graph density:")
gden(ergm_induced_net)  # unweighted density of the induced network

```

```{r}
# goodness of fit
gof(model.00)
plot(gof(model.00))
```

```{r}
#model.01

#Geometrically weighted edgewise shared partner

start.time <- Sys.time()

set.seed(0)
model.01 <- ergm(ergm_induced_net~edges+triangle)
summary(model.01)

end.time <- Sys.time()
time.taken <- end.time - start.time
time.taken

```
```{r}
print("The probability of activating a tie randomly:")
plogis(coef(model.01)[1])
print("The probability of activating a tie that creates a triangle:")
plogis(coef(model.01)[1]+coef(model.01)[2])
```


```{r}
mcmc.diagnostics(model.01)
```

```{r}
# goodness of fit
gof(model.01)
plot(gof(model.01))
```

```{r}
#model.02

#Geometrically weighted edgewise shared partner

start.time <- Sys.time()

set.seed(0)
model.02 <- ergm(ergm_induced_net~edges+gwesp)
summary(model.02)

end.time <- Sys.time()
time.taken <- end.time - start.time
time.taken

```

```{r}
print("The probability of activating a tie randomly:")
plogis(coef(model.02)[1])
print("The probability considering 2 coefficients:")
plogis(coef(model.02)[1]+coef(model.02)[2])
print("The probability considering 3 coefficients:")
plogis(coef(model.02)[1]+coef(model.02)[2]+coef(model.02)[3])
```

```{r}
mcmc.diagnostics(model.02)
```

```{r}
# goodness of fit
gof(model.02)
plot(gof(model.02))
```




```{r}
# model.03
# Geometrically weighted degree distribution
start.time <- Sys.time()

model.03 <- ergm(ergm_induced_net~edges+ gwdegree)
summary(model.03)

end.time <- Sys.time()
time.taken <- end.time - start.time
time.taken
```

```{r}
print("The probability of activating a tie randomly:")
plogis(coef(model.03)[1])
print("The probability considering 2 coefficients:")
plogis(coef(model.03)[1]+coef(model.03)[2])
print("The probability considering 3 coefficients:")
plogis(coef(model.03)[1]+coef(model.03)[2]+coef(model.03)[3])
```

```{r}
mcmc.diagnostics(model.03)
```

```{r}
# goodness of fit
summary(gof(model.03))
plot(gof(model.03))
```


```{r}
# model.04
#nodal attributes

start.time <- Sys.time()

model.04 <- ergm(ergm_induced_net~edges+
                   absdiff("Academic_Age")+ 
                   absdiff("Age")+ 
                   nodematch("Role")+ nodefactor("Role")+
                   nodematch("University")+ nodefactor("University")+
                   nodematch("Gender")+ nodefactor("University")+
                   nodematch("Faculty")+ nodefactor("Faculty"))
summary(model.04)

end.time <- Sys.time()
time.taken <- end.time - start.time
time.taken
```


```{r}
print("The probability of activating a tie randomly:")
plogis(coef(model.04)[1])
print("The probability considering 2 coefficients:")
plogis(coef(model.04)[1]+coef(model.04)[2])
print("The probability considering 3 coefficients:")
plogis(coef(model.04)[1]+coef(model.04)[2]+coef(model.04)[3])
```

```{r}
mcmc.diagnostics(model.04)
```

```{r}
# goodness of fit
gof(model.04)
plot(gof(model.04))
```

```{r}
# a simulated network from model.04
model.04.sim <- simulate(model.04,nsim=1)

```

```{r}
gplot(model.04.sim[[1]], label= model.04.sim[[1]] %v% "vertex.names", vertex.cex = model.04.sim[[1]] %v% )

```

